---
title: "SEM Model Componets"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

## Set Up & Objective 

* take a hard look at the model components of SEM 
* understand the dynamics and the sensitivity
* what is my interpretation of what is going on... 


```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)

# This requires a local installation of the developmental version of the package
SEM_root <- "~/Documents/2023/FoRTEII/SEM"
devtools::load_all(SEM_root)
inputs <- read.csv(file.path(SEM_root, "tests", "testthat", "inputs.csv"))

theme_set(theme_bw())
```

# Componet Description

# Arrhenius

Physical chemistry is the formula for temperature dependency of reaction rates, in SEM it is used to 
adjust the respiration/reaction rates that might have been observed in a lab to reflect the temperature
at a particular time. 

```{r}
lapply(0:100, FUN = arrhenius, observed.value = 0.04 * 18) %>% 
  unlist -> 
  out

data.frame(temp = 0:100, rxn = out) %>% 
  ggplot(aes(temp, rxn)) + 
  geom_line() + 
  labs(title = "Arrhenius", subtitle = "Using Photosynthesis params")

```


# Farquhar


"The Farquhar biochemical growth model (Farquhar et al., 1980) calculates photosynthesis as a function of demand and supply of CO2. The advantage with this model is that photosynthesis is regulated not only by radiation and transpiration, but also by air humidity, leaf temperature, CO2 availability and leaf nitrogen content, and the plant also experience radiation saturation at high levels of radiation. To function properly, driving variables need to be given as input to the simulation at least once an hour."^[http://www.coupmodel.com/WordDocuments/farquharapproach1.htm]

**One of the assumptions is that module needs less than hourly data!**

```{r, echo = TRUE}
# This is the molar gas constant aka molar equivalent to the Boltzmann constant,
# expressed in units of energy per temperature increment per amount of substance
R <- 8.3144621 # gas constant

# Tleaf is around 25C in units of K which is the temp the basal respiration 
# measurements are collected at. 
Tleaf <- 298 # this is 25 C from K 

Kc <- 404.9 * exp(79430 * (Tleaf - 298) / (298 * R * Tleaf)) # the min 
Ko <- 278.4 * exp(36380 * (Tleaf - 298) / (298 * R * Tleaf)) # the max?? 
Km <- Kc * (1 + 210 / Ko)
```

The photosynthesis rate is determined by if the system is light limited or CO2 limited. 

```{r, eval = FALSE}
# Params
Vcmax <- 18
Gstar <- 10 
Ci <- 400

Fparams <- c("alpha" = 0.8, "Jmax" = 1.67 * Vcmax, "Gstar" = Gstar, "Km" = Km,
             "Vcmax" = Vcmax, "Rleaf" =  0.04 * Vcmax)  
I <- inputs$PAR

a <- 0.9 ## curvature parameter
b <- -(Fparams[["alpha"]] * I + Fparams[["Jmax"]])
c <- Fparams[["alpha"]] * I * Fparams[["Jmax"]]
J <- (-b -sqrt(b^2-4*a*c))/ (2*a)
aJ_400 <- J*(Ci-Fparams[["Gstar"]])/(4*Ci+8*Fparams[["Gstar"]])    ## electron transport limited without covariates
aC <- Fparams[["Vcmax"]]*(Ci-Fparams[["Gstar"]])/(Ci + Fparams[["Km"]])  

# okay so CO2 does range from 305.3 429.8 but they leave it alone in SEM so much be assumed to be constant
# so will probably need to just leave it alone

df <- data.frame(time = inputs$time, 
                 par = inputs$PAR, 
                 aJ = aJ - Fparams[["Rleaf"]], 
                 aC = aC -  Fparams[["Rleaf"]])

ggplot(data = df) + 
  geom_line(aes(par, aJ)) + 
  geom_line(aes(par, aC)) + 
  labs(title = "Photosynthesis", 
       y = "Photosynthesis", 
       x = "PAR")

```
# Ballberry

This function dete

# Hydrology 

Patch hydrology occurs in two steps 

* soil to atmosphere evaporation (environmental conditions)
* plant water use (driven by photosynthesis) 


```{r}
# Conversion factors and constants 
conversion_factor <- (1/1000) # kg/m2/s to m/s
timestep <- 1800
# Constants
rho <- 1.15           # density of air, kg/m3 
P <- 101.325          # average atm pressure (kPa)
R <- 8.3144621        # ideal gas constant in J/K/mol

EVAP = min(X[["soil_water"]]/timestep, rho * params$gevap * (0.622 * inputs[["VPD"]] / P) * conversion_factor)  
```



```{r, eval = FALSE}
gevap <- 0.005  
rho *  gevap

x <- (0.622 * inputs[["VPD"]] / P) * conversion_factor * rho
range(inputs[["VPD"]] / P)


```

